Slitasje og kjørespor

### 2023-01-18 DRAFT_slitasjeOgKjørespor.Rmd
feat: HIF per region

```{r}
regions <- sf::read_sf("data/regions.shp")
```

@@

### Homogeneous impact areas
I want to use the [homogeneous impact areas]() (HIA) to define smaller regions into which I can extrapolate the indicator values.
This data is generated by discretizing the Norwegian Infrastructure Index. I refer to the ordinal values of the
four impact area classes as their *Human Impact Factor*.

@@

I also want to split these four HIA classes based on their region.

<!-- Add function to ea_homogeneous_area() to find intersection with a decond delinimator -->

>>>>

```{r}
regions <- sf::read_sf("data/regions.shp", options = "ENCODING=UTF8")
unique(regions$region)
```

@@

### Homogeneous impact areas
I want to use the [homogeneous impact areas]() (HIA) to define smaller regions into which I can extrapolate the indicator values.
This data is generated by discretizing the Norwegian Infrastructure Index. I refer to the ordinal values of the
four impact area classes as their *Human Impact Factor* (HIF).

@@

I also want to split these four HIA classes based on their region. To do this I need the two layers to have the same crs
```{r}
st_crs(HIA) == st_crs(regions)
```

```{r}
HIA <- st_transform(HIA, 25833)
```

```{r}
st_crs(HIA) == st_crs(regions)
```

Since the two layers are completely overlapping, we can get the intersections
```{r, eval=F}
HIA_regions <- eaTools::ea_homogeneous_area(HIA,
                             regions,
                             keep1 = "infrastructureIndex",
                             keep2 = "region")
saveRDS(HIA_regions, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/HIA_regions.rds")
```

```{r, echo=F}
# silently read cached file
HIA_regions <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/cache/HIA_regions.rds")
```

I now have 20 unique areas that I will, given there is data, extrapolate indicator values over: 
```{r}
HIA_regions <- HIA_regions %>%
  mutate(region_HIF = paste(region, infrastructureIndex))
length(unique(HIA_regions$region_HIF))
```

Let's calculate the areas of these polygons and compare the HIF in the five regions. 
```{r HIF-region, fig.cap="Stacked barplot showing the distribution of human impact factor across five regions in Norway."}
HIA_regions$area <- sf::st_area(HIA_regions)

temp <- as.data.frame(HIA_regions) %>%
  group_by(region, infrastructureIndex) %>%
  summarise(area = mean(area))

ggplot(temp, aes(x = region, y = area, fill = factor(infrastructureIndex)))+
  geom_bar(position = "stack", stat = "identity")
```

The figure above shows that Nord-Norge for example, has a lot of relatovely untouched areas, and that Østlandet has the highets proportion of impacted areas.
This is expected.
